---
title: "nkelley_OriginalHomeworkCode_04"
author: "maekell98"
date: "10/20/2021"
output: html_document
---

[1] Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data

```{r Z.prop.test function}

Z.prop.test <- function(p1, n1, p2 = NULL, n2 = NULL, p0, alternative = "two.sided", conf.level = 0.95) 
#specifies arguments
{
  pstar <- ((p2*n2)+(p1*n1))/(n1 + n2)
#pstar is the pooled proportion for both samples
  
  Z <- (p2-p1-p0)/(sqrt(pstar(1-pstar)(1/n1 + 1/n2)))
  return(Z)
#Z test statistic for two-sample Z test
  
  P <- 1 - pnorm(Z, lower.tail = TRUE) + pnorm(Z, lower.tail = FALSE)
  return(P)
#I believe this p value is for if the two proportions differ significantly from one another, but I need to look into it more.
  
  lower <- p1 - qnorm(0.975) * sqrt(p1 * (1 - p1)/30)
upper <- p1 + qnorm(0.975) * sqrt(p1 * (1 - p1)/30)
CI <- c(lower, upper)
  return(CI)
#95% confidence interval based on normal distribution  

  if(alternative == "less"){
    P <- pnorm(Z, lower.tail = TRUE)
    P
  } 
#here specify what should be done if two sample (if p1 smaller aka alternative = "less"/larger aka alternative = "greater" than p2 -- same as x and y in function t.test())
##this calculates one-sided p value of p1<p2

  if(alternative == "greater"){
    P <- pnorm(Z, lower.tail = FALSE)
    P
} 
#this calcuates one-sided p value of p1>p2

  if(is.null(p2) || is.null(n2)){
    
    Z <- (p1-p0)/(sqrt(p0(1-p0)/n1))
    return(Z)
#Z statistic for one-sample Z test    
    
    P <- pnorm(Z, two.sided = TRUE)
  return(P)
#p value associated with Z for one-sample Z test
    
     if(alternative == "greater"){P <- pnorm(Z, lower.tail = FALSE)
  return(P)}
    
    if(alternative == "less"){P <- pnorm(Z, lower.tail = TRUE)
  return(P)}
#p values for greater and less than alternatives  
    
    lower <- p1 - qnorm(0.975) * sqrt(p1 * (1 - p1)/30)
    upper <- p1 + qnorm(0.975) * sqrt(p1 * (1 - p1)/30)
CI <- c(lower, upper)
  return(CI)
#CI for one-sample Z test
 
}
#here if p2 or n2 or both is NULL, perform one-sample Z test. "||" means [or](http://applied-r.com/conditionals-in-r/)

  if((n1 * p1 > 5) == FALSE || (n1 * (1-p1) > 5) == FALSE || (n2 * p2 > 5) == FALSE || (n2 * (1-p2) > 5) == FALSE) {
    warning("normal distribution invalid")} 
#here specify that if (\(n * p > 5\) and \(n * (1-p) > 5\)) untrue, run function but print error message that assumptions aren't met. Included for p2/n2 as well. Not sure if these should be separated somehow. 

#return Z (test statistic), P (appropriate p value), CI (two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test; calc symmetric CIs based on quantiles of normal distribution rather than single-limit confidence bounds!)
  

} # end of function

```

```{r load}
install.packages("tidyverse")

install.packages("ggplot2")

library(curl)

library(ggplot2)

library(tidyverse)
```

```{r load regression packages}

install.packages("gridExtra")
install.packages("manipulate")
install.packages("lmodel2")
```

```{r library}
library(gridExtra)
library(manipulate)
library(lmodel2)
```

[2] The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size):

```{r load in Kamilar Cooper data}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall21/KamilarAndCooperData.csv")

d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
#creates data frame from raw csv file
```

1) Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).

```{r regression and equation}

m <- lm(MaxLongevity_m ~ Brain_Size_Species_Mean, data = d)
m

t <- coef(summary(m))
t <- data.frame(unlist(t))
colnames(t) <- c("Est", "SE", "t", "p")
t

eq <- function(x,y) {
  m <- lm(y ~ x)
  as.character(
    as.expression(
      substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                list(a = format(coef(m)[1], digits = 4),
                b = format(coef(m)[2], digits = 4),
                r2 = format(summary(m)$r.squared, digits = 3)))
    )
  )
}
```
eq() from https://www.roelpeters.be/how-to-add-a-regression-equation-and-r-squared-in-ggplot2/

```{r ggplot linear regressions}
g <- ggplot(data = d, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m), xlab = "Brain Size Species Mean", ylab = "Max Longevity", main = "Brain Size vs Longevity")
g <- g + geom_point()
g <- g + geom_smooth(method = "lm", formula = y ~ x)
g <- g + geom_text(x = 200, y = 775, label = eq(d$MaxLongevity_m, d$Brain_Size_Species_Mean), parse = TRUE)
g
```

```{r model II regression analysis}
mII <- lmodel2(MaxLongevity_m ~ Brain_Size_Species_Mean, data = d, range.y = "relative", range.x = "relative",
    nperm = 1000)
mII

plot(mII, "OLS", xlab = "Brain Size Species Mean", ylab = "Max Longevity", main = "Brain Size vs Longevity")


g <- ggplot(data = d, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m))
g <- g + geom_point()
g <- g + geom_smooth(method = "lmodel2", formula = y ~ x)
g <- g + geom_text(x = 200, y = 700, label = eq(d$MaxLongevity_m, d$Brain_Size_Species_Mean), parse = TRUE)
g
```
I wasn't sure if I needed to use ModelII analysis, but when I tried I wasn't sure how to add the linear regression line. 

2) Identify and interpret the point estimate of the slope (\(\beta_1\)), as well as the outcome of the test associated with the hypotheses H0: \(\beta_1\) = 0; HA: \(\beta_1\) ≠ 0. Also, find a 90 percent CI for the slope (\(\beta_1\)) parameter.

```{r id and interpret slope and test, also CI}

beta1 <- t$Est[2]
beta1

```

3) Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.
```{r ad 90% CI and legend}

```

4) Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?
```{r 90% PI}

```

5) Looking at your two models, which do you think is better? Why?

